---
title: "Check-In Presentation Materials"
author: "Ethan Rodriguez-Shah"
date: "10/27/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```

```{r, include = FALSE}
## Set working directory before running code
getwd <- "~/Google Drive/6. ECON 181 - Expository Data Analysis w: R/Final Project"
setwd(getwd)
```


# Introduction

With the population changes and shifting demographics in the United States capital brought on in the aftermath of the 2008 economic crisis, real estate is an incredibly important business, coveted by real estate agencies and developers and peddled by the District government. Many different factors can affect property values, including number of bedrooms, number of bathrooms, style of kitchen, square footage, and style of flooring, among others. A University of Michigan study conducted by the Department of Horticulture found that even plant material, size, and even design can affect the value of a home (Barton et al). Through my research, I will be examining the effect that proximity to Metro stations has on property prices in the Washington, D.C., Metropolitan Area. In examining property data, I expect to find that Metro station proximity and property prices will have negative correlation; that is, property prices will increase the nearer a property is to a station.

# Data Description

My research will be conducted using two datasets: the first, station_locations, is a list of the ninety-one Metro stations in the WMATA system, provided by Federal Reserve economist Neil Bhutta; the second, redfin_data, consists of all the properties in the Washington Metropolitan Area listed by Redfin, a Seattle-based residential real estate company, as sold in the three months prior to October 19, 2017. I compiled the panel data frame redfin_data manually from the Redfin website. Redfin only allows downloads of data sets with the information of 350 or less properties at one time, so I compiled the redfin_data by reading in many datasets and joining them by the ADDRESS variable, so that no single property would be included in the data twice. The station_locations data frame will be useful for answering my research question because it contains location data for all WMATA stations, and the redfin_data data frame will be useful because it includes not only location data but also data on other property characteristics like number of bedrooms, number of bathrooms, and square footage. This information should be sufficient to allow me to test my hypothesis.

## Key economic variables

This research centers around only one completely dependent variable, two completely independent variable, and many other variables that may shift between being dependent and independent. The dependent variable is property price. The independent variables are date sold and distance from nearest Metro station. The flexible variables are as follows:

    * Property type
    * Number of bedrooms
    * Number of bathrooms
    * Square footage
    * Days on market
    
## Data loading and cleaning

Because of the nature in which I had to collect the data, my loading and cleaning process was very long.

### Loading necessary libraries

```{r}
## Necessary libraries
#install.packages("RCurl")
library(RCurl)
library(dplyr)
library(geosphere)
library(stringr)
library(lubridate)
```

### Read in data

